---
title: "Income Inequality Before and After Taxes"
author: "Aditya Arya"
format: html
execute:
  message: false
  warning: false
---

## Question

How do pre-tax and post-tax income inequality compare across countries in 2020?

The Gini coefficient measures income inequality from 0 to 1. Higher numbers mean more inequality. The data come from [TidyTuesday's income inequality dataset](https://github.com/rfordatascience/tidytuesday/tree/main/data/2025/2025-08-05).

In the file used here, each row is a country (or region) in one year. I use `Year` to select 2020, `gini_mi_eq` for the pre-tax measure, and `gini_dhi_eq` for the post-tax measure. Both measures adjust income for household size. "Pre-tax" and "post-tax" are short names for income before and after taxes **and benefits**.

```{r}
library(tidyverse)

income_inequality <- read_csv(
  "https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-08-05/income_inequality_processed.csv"
)
```

## Choose the observations

I use 2020 because the dataset has both Gini measures for 19 countries in that year. There are 31 rows for 2020; 12 have no pre-tax Gini value, while none are missing the post-tax value. I remove rows missing either measure. This means the graph compares only countries with both values available, not every country in the world.

```{r}
inequality_2020 <- income_inequality |>
  filter(
    Year == 2020,
    !is.na(gini_mi_eq),
    !is.na(gini_dhi_eq)
  )
```

## Make a graph

Each point represents one country. The horizontal position shows its pre-tax Gini coefficient, and the vertical position shows its post-tax Gini coefficient.

```{r}
ggplot(inequality_2020, aes(x = gini_mi_eq, y = gini_dhi_eq)) +
  geom_point() +
  labs(
    title = "Income inequality before and after taxes in 2020",
    x = "Pre-tax Gini coefficient",
    y = "Post-tax Gini coefficient",
    caption = "Source: Our World in Data via TidyTuesday"
  )
```

In this dataset, the post-tax Gini coefficient is lower than the pre-tax coefficient for all 19 countries shown. This is a comparison of the measures, not proof that a particular policy caused the difference.

## Sources

- TidyTuesday. (2025, August 5). [Income Inequality Before and After Taxes](https://github.com/rfordatascience/tidytuesday/tree/main/data/2025/2025-08-05) (data and variable descriptions).
- Hasell, J. (2023, July 3). [Income inequality before and after taxes: how much do countries redistribute income?](https://ourworldindata.org/income-inequality-before-and-after-taxes) *Our World in Data* (original data processing and explanation).
- Luxembourg Income Study. (2025). [LIS Database](https://www.lisdatacenter.org/our-data/lis-database/) (underlying household-income data, as credited by TidyTuesday and Our World in Data).
