---
title: "Income Inequality Before and After Taxes"
author: "Aditya Arya"
format: html
---

## Question

How do pre-tax and post-tax income inequality compare across countries in 2020?

The Gini coefficient measures income inequality from 0 to 1. Higher numbers mean more inequality. The data come from [TidyTuesday's income inequality dataset](https://github.com/rfordatascience/tidytuesday/tree/main/data/2025/2025-08-05).

```{r}
#| message: false
#| warning: false

library(tidyverse)

income_inequality <- read_csv(
  "https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-08-05/income_inequality_processed.csv"
)
```

## Choose the observations

I use 2020 because the dataset has both Gini measures for 19 countries in that year. I remove rows missing either measure.

```{r}
inequality_2020 <- income_inequality |>
  filter(
    Year == 2020,
    !is.na(gini_mi_eq),
    !is.na(gini_dhi_eq)
  )
```

## Make a graph

Each point represents one country. The horizontal position shows its pre-tax Gini coefficient, and the vertical position shows its post-tax Gini coefficient.

```{r}
ggplot(inequality_2020, aes(x = gini_mi_eq, y = gini_dhi_eq)) +
  geom_point() +
  labs(
    title = "Income inequality before and after taxes in 2020",
    x = "Pre-tax Gini coefficient",
    y = "Post-tax Gini coefficient",
    caption = "Source: Our World in Data via TidyTuesday"
  )
```

In this dataset, the post-tax Gini coefficient is lower than the pre-tax coefficient for all 19 countries shown. This is a comparison of the measures, not proof that a particular policy caused the difference.
