---
title: "Income Inequality Before and After Taxes"
author: "Aditya Arya"
format: html
---

## Introduction

The Gini coefficient measures income inequality from 0 to 1, with higher values representing greater inequality. This analysis examines which countries experience the largest reduction in inequality after taxes and government benefits.

The data come from the [TidyTuesday Income Inequality dataset](https://github.com/rfordatascience/tidytuesday/tree/main/data/2025/2025-08-05).

```{r}
#| message: false
#| warning: false

library(tidyverse)

income_inequality <- readr::read_csv(
  "https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-08-05/income_inequality_processed.csv"
)
```

## Preparing the data

For each country, I use its most recent observation and calculate the difference between its pre-tax and post-tax Gini coefficients.

```{r}
latest_inequality <- income_inequality |>
  filter(
    !is.na(Code),
    !is.na(gini_mi_eq),
    !is.na(gini_dhi_eq)
  ) |>
  group_by(Entity) |>
  slice_max(Year, n = 1, with_ties = FALSE) |>
  ungroup() |>
  mutate(
    inequality_reduction = gini_mi_eq - gini_dhi_eq
  )

top_reducers <- latest_inequality |>
  slice_max(inequality_reduction, n = 15)
```

## Which countries reduce inequality the most?

```{r}
ggplot(
  top_reducers,
  aes(
    x = inequality_reduction,
    y = reorder(Entity, inequality_reduction)
  )
) +
  geom_col(fill = "#2C7FB8") +
  labs(
    title = "Countries with the largest reductions in income inequality",
    subtitle = "Difference between pre-tax and post-tax Gini coefficients",
    x = "Reduction in Gini coefficient",
    y = NULL,
    caption = "Source: Our World in Data via TidyTuesday"
  ) +
  theme_minimal()
```

The graph compares each country's most recent available observation. A larger difference indicates that taxes and government benefits are associated with a greater reduction in measured income inequality.
